---
title: "Titanic: Machine Learning from Disaster"
output: html_notebook
---

Introduction
------------
The sinking of the RMS Titanic is one of the most infamous shipwrecks in history.  On April 15, 1912, during her maiden voyage, the Titanic sank after colliding with an iceberg, killing 1502 out of 2224 passengers and crew. This sensational tragedy shocked the international community and led to better safety regulations for ships.

One of the reasons that the shipwreck led to such loss of life was that there were not enough lifeboats for the passengers and crew. Although there was some element of luck involved in surviving the sinking, some groups of people were more likely to survive than others, such as women, children, and the upper-class.

In this challenge, we ask you to complete the analysis of what sorts of people were likely to survive. In particular, we ask you to apply the tools of machine learning to predict which passengers survived the tragedy.

Overview
--------
The data has been split into two groups:

training set (train.csv)
test set (test.csv)
The training set should be used to build your machine learning models. For the training set, we provide the outcome (also known as the “ground truth”) for each passenger. Your model will be based on “features” like passengers’ gender and class. You can also use feature engineering to create new features.

The test set should be used to see how well your model performs on unseen data. For the test set, we do not provide the ground truth for each passenger. It is your job to predict these outcomes. For each passenger in the test set, use the model you trained to predict whether or not they survived the sinking of the Titanic.

We also include gender_submission.csv, a set of predictions that assume all and only female passengers survive, as an example of what a submission file should look like.

Data Dictionary
---------------
  Variable	    Definition	      Key
  survival	    Survival	        0 = No, 1 = Yes
  pclass	      Ticket class	    1 = 1st, 2 = 2nd, 3 = 3rd
  sex	          Sex	
  Age	          Age in years	
  sibsp	        # of siblings / spouses aboard the Titanic	
  parch	        # of parents / children aboard the Titanic	
  ticket	      Ticket number	
  fare	        Passenger fare	
  cabin	        Cabin number	
  embarked	    Port of Embarkation	C = Cherbourg, Q = Queenstown, S = Southampton

Variable Notes
--------------
  pclass: A proxy for socio-economic status (SES)
  1st     = Upper
  2nd     = Middle  
  3rd     = Lower

  age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

  sibsp: The dataset defines family relations in this way...
  Sibling = brother, sister, stepbrother, stepsister
  Spouse = husband, wife (mistresses and fiancés were ignored)

  parch: The dataset defines family relations in this way...
  Parent = mother, father
  Child = daughter, son, stepdaughter, stepson
  Some children travelled only with a nanny, therefore parch=0 for them.
  
Load Libraries and Data
-----------------------

```{r}
# Libraries
library("ggplot2")
library("ggthemes")

#testData <- read.csv("./Data/test.csv", header = T)
factTable <- read.csv("./Data/train.csv", header = T)
```

Segmentation
------------

My interest in in the probablitiy of survival based on passenger attributes. Passenger attributes which may influence the probablity of survival is age, sex, the class they were seated in, family size, and port of departure.

The intent is to use this information for Information Gain(IG)
```{r}

# Begin Segmentation of population
factTable$AgeGroup[factTable$Age <= 18] <- 'CHILD'
factTable$AgeGroup[factTable$Age > 18 & factTable$Age < 65] <- 'ADULT'
factTable$AgeGroup[factTable$Age >= 65] <- 'SENIOR'

factTable$FSize <- factTable$SibSp + factTable$Parch + 1
factTable$FamilySize[factTable$FSize == 1] <- 'SINGLE'
factTable$FamilySize[factTable$FSize > 1 & factTable$FSize <= 4] <- 'SMALL'
factTable$FamilySize[factTable$FSize > 4 & factTable$FSize <= 6] <- 'MEDIUM'
factTable$FamilySize[factTable$FSize > 6] <- 'LARGE'
```

Measures
--------

percenatge of the segmentaiton rleated to the total amount of passengers
```{r}
# Percentage values used for information gain calculation
TotalPassengers <- nrow(factTable)

# Age Group
percentageOfChildren <- length(which(factTable$AgeGroup == 'CHILD'))/TotalPassengers
percentageOfAdults <- length(which(factTable$AgeGroup == 'ADULT'))/TotalPassengers
percentageOfSeniors <- length(which(factTable$AgeGroup == 'SENIOR'))/TotalPassengers

# Family Size
percentageOfSingles <- length(which(factTable$FamilySize == 'SINGLE'))/TotalPassengers
percentageOfSmallFamilies <- length(which(factTable$FamilySize == 'SMALL'))/TotalPassengers
percentageOfMediumFamilies <- length(which(factTable$FamilySize == 'MEDIUM'))/TotalPassengers
percentageOfLargeFamilies <- length(which(factTable$FamilySize == 'LARGE'))/TotalPassengers

# Embark
percentageFromC <- length(which(factTable$Embarked == 'C'))/TotalPassengers
percentageFromQ <- length(which(factTable$Embarked == 'Q'))/TotalPassengers
percentageFromS <- length(which(factTable$Embarked == 'S'))/TotalPassengers

# Gender
percentageFemales <- length(which(factTable$Sex == 'female'))/TotalPassengers
percentageMales <- length(which(factTable$Sex == 'male'))/TotalPassengers

# Class
percentageClass1 <- length(which(factTable$Pclass == 1))/TotalPassengers
percentageClass2 <- length(which(factTable$Pclass == 2))/TotalPassengers
percentageClass3 <- length(which(factTable$Pclass == 3))/TotalPassengers


# Entropy by Age Group
total <- length(which(factTable$AgeGroup == 'CHILD'))
survived <- length(which(factTable$AgeGroup == 'CHILD' & factTable$Survived == 1))
dead <- length(which(factTable$AgeGroup == 'CHILD' & factTable$Survived == 1))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1) + p2*log2(p2))
factTable$EntropyByAgeGroup[factTable$AgeGroup == 'CHILD'] <- entropy
factTable$PercentageByAgeGroup[factTable$AgeGroup == 'CHILD'] <- percentageOfChildren

total <- length(which(factTable$AgeGroup == 'SENIOR'))
survived <- length(which(factTable$AgeGroup == 'SENIOR' & factTable$Survived == 1))
dead <- length(which(factTable$AgeGroup == 'SENIOR' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1) + p2*log2(p2))
factTable$EntropyByAgeGroup[factTable$AgeGroup == 'SENIOR'] <- entropy
factTable$PercentageByAgeGroup[factTable$AgeGroup == 'SENIOR'] <- percentageOfSeniors

total <- length(which(factTable$AgeGroup == 'ADULT'))
survived <- length(which(factTable$AgeGroup == 'ADULT' & factTable$Survived == 1))
dead <- length(which(factTable$AgeGroup == 'ADULT' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1) + p2*log2(p2))
factTable$PercentageByAgeGroup[factTable$AgeGroup == 'ADULT'] <- percentageOfAdults
factTable$EntropyByAgeGroup[factTable$AgeGroup == 'ADULT'] <- entropy


# Entropy by Embark
total <- length(which(factTable$Embarked == 'C'))
survived <- length(which(factTable$Embarked == 'C' & factTable$Survived == 1))
dead <- length(which(factTable$Embarked == 'C' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByEmbark[factTable$Embarked == 'C'] <- percentageFromC
factTable$EntropyByEmbark[factTable$Embarked == 'C'] <- entropy

total <- length(which(factTable$Embarked == 'S'))
survived <- length(which(factTable$Embarked == 'Q' & factTable$Survived == 1))
dead <- length(which(factTable$Embarked == 'Q' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByEmbark[factTable$Embarked == 'Q'] <- percentageFromQ
factTable$EntropyByEmbark[factTable$Embarked == 'Q'] <- entropy

total <- length(which(factTable$Embarked == 'S'))
survived <- length(which(factTable$Embarked == 'S' & factTable$Survived == 1))
dead <- length(which(factTable$Embarked == 'S' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByEmbark[factTable$Embarked == 'S'] <- percentageFromS
factTable$EntropyByEmbark[factTable$Embarked == 'S'] <- entropy


# Entropy by Class
total <- length(which(factTable$Pclass == 1))
survived <- length(which(factTable$Pclass == 1 & factTable$Survived == 1))
dead <- length(which(factTable$Pclass == 1 & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByClass[factTable$Pclass == 1] <- percentageClass1
factTable$EntropyByClass[factTable$Pclass == 1] <- entropy

total <- length(which(factTable$Pclass == 2))
survived <- length(which(factTable$Pclass == 2 & factTable$Survived == 1))
dead <- length(which(factTable$Pclass == 2 & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByClass[factTable$Pclass == 2] <- percentageClass2
factTable$EntropyByClass[factTable$Pclass == 2] <- entropy

total <- length(which(factTable$Pclass == 3))
survived <- length(which(factTable$Pclass == 3 & factTable$Survived == 1))
dead <- length(which(factTable$Pclass == 3 & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByClass[factTable$Pclass == 3] <- percentageClass3
factTable$EntropyByClass[factTable$Pclass == 3] <- entropy

# Entropy by Gender
total <- length(which(factTable$Sex == 'female'))
survived <- length(which(factTable$Sex == 'female' & factTable$Survived == 1))
dead <- length(which(factTable$Sex == 'female' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByGender[factTable$Sex == 'female'] <- percentageFemales
factTable$EntropyByGender[factTable$Sex == 'female'] <- entropy

total <- length(which(factTable$Sex == 'male'))
survived <- length(which(factTable$Sex == 'male' & factTable$Survived == 1))
dead <- length(which(factTable$Sex == 'male' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByGender[factTable$Sex == 'male'] <- percentageMales
factTable$EntropyByGender[factTable$Sex == 'male'] <- entropy


# Entropy by Family size
total <- length(which(factTable$FamilySize == 'SINGLE'))
survived <- length(which(factTable$FamilySize == 'SINGLE' & factTable$Survived == 1))
dead <- length(which(factTable$FamilySize == 'SINGLE' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByFamilySize[factTable$FamilySize == 'SINGLE'] <- percentageOfSingles
factTable$EntropyByFamilySize[factTable$FamilySize == 'SINGLE'] <- entropy

total <- length(which(factTable$FamilySize == 'SMALL'))
survived <- length(which(factTable$FamilySize == 'SMALL' & factTable$Survived == 1))
dead <- length(which(factTable$FamilySize == 'SMALL' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByFamilySize[factTable$FamilySize == 'SMALL'] <- percentageOfSmallFamilies
factTable$EntropyByFamilySize[factTable$FamilySize == 'SMALL'] <- entropy

total <- length(which(factTable$FamilySize == 'MEDIUM'))
survived <- length(which(factTable$FamilySize == 'MEDIUM' & factTable$Survived == 1))
dead <- length(which(factTable$FamilySize == 'MEDIUM' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByFamilySize[factTable$FamilySize == 'MEDIUM'] <- percentageOfMediumFamilies
factTable$EntropyByFamilySize[factTable$FamilySize == 'MEDIUM'] <- entropy

total <- length(which(factTable$FamilySize == 'LARGE'))
survived <- length(which(factTable$FamilySize == 'LARGE' & factTable$Survived == 1))
dead <- length(which(factTable$FamilySize == 'LARGE' & factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1)+p2*log2(p2))
factTable$PercentageByFamilySize[factTable$FamilySize == 'LARGE'] <- percentageOfLargeFamilies
factTable$EntropyByFamilySize[factTable$Sex == 'LARGE'] <- entropy


#Overall Survival
total <- nrow(factTable)
survived = length(which(factTable$Survived == 1))
dead = length(which(factTable$Survived == 0))
p1 <- survived/total
p2 <- dead/total
entropy <- -(p1*log2(p1) + p2*log2(p2))
factTable$EntropyParentSet <- entropy


# Remove Objects
rm(p1)
rm(p2)
rm(survived)
rm(dead)
rm(entropy)
rm(total)
```


Subsetting and Graphing
-----------------------
```{r}
AgeGroupSubset <- na.omit(factTable[which(!duplicated(factTable$AgeGroup)), c('AgeGroup', 'PercentageByAgeGroup', 'EntropyByAgeGroup')])

FamilySizeSubset <- na.omit(factTable[which(!duplicated(factTable$FamilySize)), c('FamilySize', 'PercentageByFamilySize', 'EntropyByFamilySize')])

GenderSubset <- na.omit(factTable[which(!duplicated(factTable$Sex)), c('Sex', 'PercentageByGender', 'EntropyByGender')])

ClassSubset <- na.omit(factTable[which(!duplicated(factTable$Pclass)), c('Pclass', 'PercentageByClass', 'EntropyByClass')])

EmbarkSubset <- na.omit(factTable[which(!duplicated(factTable$Embarked)), c('Embarked', 'PercentageByEmbark', 'EntropyByEmbark')])


AgeGroupSubset$InformationGain <- AgeGroupSubset$PercentageByAgeGroup * AgeGroupSubset$EntropyByAgeGroup

FamilySizeSubset$InformationGain <- FamilySizeSubset$PercentageByFamilySize * FamilySizeSubset$EntropyByFamilySize

GenderSubset$InformationGain <- GenderSubset$PercentageByGender * GenderSubset$EntropyByGender

ClassSubset$InformationGain <- ClassSubset$PercentageByClass * ClassSubset$EntropyByClass

EmbarkSubset$InformationGain <- EmbarkSubset$PercentageByEmbark * EmbarkSubset$EntropyByEmbark

# View(AgeGroupSubset)
# View(FamilySizeSubset)
# View(GenderSubset)
# View(ClassSubset)
# View(EmbarkSubset)

IG_AgeGroup <- sum(AgeGroupSubset$InformationGain)
IG_FamilySize <- sum(FamilySizeSubset$InformationGain)
IG_Gender <- sum(GenderSubset$InformationGain)
IG_Class <- sum(ClassSubset$InformationGain)
IG_Embark <- sum(EmbarkSubset$InformationGain)

FinalEntropyParent <- factTable$EntropyParentSet - IG_AgeGroup - IG_FamilySize - IG_Gender - IG_Class - IG_Embark

factTable$EntropyParentSetAfter <- FinalEntropyParent

View(factTable)
```





